### 2024-11-06 docs/outside_parameters.tex
Added docs

\begin{tabular}{@{}llll@{}}
    \toprule
    Parameter 			& 		 			& Value  	& Source \\	\midrule
    Period Length		& --				& 2 years   & PSID Frequency \\
    Discount Factor$^\dagger$		& $\beta$			& \parbeta		& Literature (see text)  \\
    Risk Aversion		& $\gamma$			& \pargamma		& Literature (see text)  \\
    Expenditure Share Housing & $\xi$		& \parxi        & Literature (see text) \\
    Rental Price$^\dagger$ 		& $q$				& \parq 		& \cite{Davis2008}  \\
    Deprecation$^\dagger$ 		& $\delta$			& \pardelta		& \cite{Harding2007} \\			
    Risk-free Rate$^\dagger$		& $r^f$ 			& \parrf 	    & Literautre (see text) \\
    Mortgage premium$^\dagger$	& $r^m$ 			& \parrm 	    & Literature (see text) \\
    PMI premium$^\dagger$         & $r^{pmi}$         & \parrpmi      & \cite{goodman2017sixty} \\
    PMI limit           & $\overline{PMI}$   & \parPMIlim     & \cite{goodman2017sixty} \\ 
    House Price          & $p$		& \parprice      & Data (see text) \\
    Max Loan-to-Value   & LTV 			& \parLTV  & Literature/data (see text) \\
    Selling \& Buying Cost		& ($m_s,m_b$) 			& (\parms,\parmb) &  \cite{Yang2009} \\ 
    Rental Size & $h_r$ & \parhr & Normalization \\
    Deterministic Income & $ l_a$				& Fig. \ref{fig:y_a}		& PSID \\
    Productivity Shocks for Kids & $y,\Pi(y'|y)$ & Fig. \ref{fig:y_ia},\ref{fig:Piv}      & PSID \\
    Initial Distribution & $F(x_{53},v_{53}$)	& Fig. \ref{fig:inidistr}		& PSID \\			
    \bottomrule			
\end{tabular}