### 2025-05-07 tabfig/postest/BW_mostrecent.tex

\begin{tabular}{l rrr rrr}
\toprule & \multicolumn{3}{c}{White} & \multicolumn{3}{c}{Black} \\ \cmidrule(lr){2-4}\cmidrule(lr){5-7}
\multicolumn{1}{c}{Moment} & \multicolumn{1}{c}{Data} & \multicolumn{1}{c}{Altr.} & \multicolumn{1}{c}{No Altr.} & \multicolumn{1}{c}{Data} & \multicolumn{1}{c}{Altr.} & \multicolumn{1}{c}{No Altr.}\\
\midrule
\;Owner (25-44) & 0.53 & 0.52 & 0.39 & 0.26 & 0.24 & 0.22\\
\;Rent / Income (25-44) & 0.22 & 0.23 & 0.21 & 0.25 & 0.27 & 0.27\\
\;Wealth at Purchase (25-44) & 43.69 & 33.45 & 46.65 & 17.17 & 51.73 & 59.28\\
\;Transfer Rate (55-74) & 0.27 & 0.24 & 0.00 & 0.14 & 0.23 & 0.00\\
\bottomrule
\end{tabular}